### 2022-03-28 paper/article.Rmd
first commit

---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.align = "center",
  fig.path = "plots/"
)
source("../misc/functions.R")

```

## What Drives Protest Coverage? {#sec:factors}

Most of the evidence we have regarding the factors which shape the conscious or unconscious choice of frames by the news media when covering protest comes from the strands of literature mentioned in Section ref(sec:paradigm).
As stated, there is a long tradition in the sociology of news production and other fields that try to uncover principal factors that shape the news.
However, what became clear is that the task is not an easy one as "the list of possible variables is almost endless" [@beerEcquidNoviSearch2010, p. 189].
@shoemakerMediatingMessage21st2014, as arguably one of the most prominent examples of the field, propose to divide factors into five distinct levels, ordered from micro to macro:
The *individual* level, which describes the characteristics of the individual journalist or editor;
the *routines* level, which describes the immediate constraining and enabling structures and routines within journalists work;
the *organisation* level, which comprises the larger context of the routinised activities such as the goals, policies and structure of an organisation;
the *social institution* level, which comprises influences on media types, such as the press, or other subsets of the media landscape;
and the *social system* level, which describes influences on content from the system as a whole.
What this thesis is most interested in are factors on the level of routines, organisations and the broader social system, which are thought to undergo several changes in the selected time frame.

Based on the existing literature, both the broader sociology of news and previous studies about protest coverage, I formulate theoretical expectations in this section regarding the third research question:

> `r hyp("RQ3")`

While the concept of testing factors which condition protest coverage is not new [e.g., @boyleAdherenceProtestParadigm2012; @kilgoHierarchy2019; @leeTriggeringProtestParadigm2014], the set of variables presented in this section is novel and, to my knowledge, the most complete so far.
To make discussion of the factors easier, I follow @shoemakerMediatingMessage21st2014 by dividing them into levels, ordered from micro to macro.
However, the three levels I use are different as they are closer tailored to the specific topic at hand --- finding the influences on how *protest* is reported.
Specifically, the factors taken into account here are divided into event-level, outlet-level and factors connected to time.

Note that at this stage, it is unknown which frames the analysis will reveal, making it impossible to specify which frames might correlate with which factors.
Many previous studies worked with the dichotomy of coverage adhering to the *protest paradigm* and coverage that is more supportive of a protest goal --- which I will call *delegitimising* and *legitimising* coverage in the following.
As it is expected that most frames can be placed within this broad dichotomy, the hypotheses in this section capture the expected relation between factors and a potential set of frames that are, most likely, legitimising or delegitimising.
In other words, while the analysis allows the revelation of more fine-grained frames, this section will discuss the theoretical expectations for relationships between the factors with legitimising and/or delegitimising coverage of protest.
However, the dichotomy of legitimising and delegitimising coverage is unlikely to tell the entire story.
What aspects of an event do news highlight and which do they omit specifically?
Depending on the specific framing, the degree to which reports legitimise or delegitimise a protest cause could vary substantially.
So while the described dichotomy is useful to discuss factors that drive reporting, the ultimate goal is to measure the influence these factors might have on the usage of specific frames.

### Event-Level Factors

#### Protest Goals

Several studies document that the goals or causes voiced by protesters shape the coverage these events receive.
These studies are based on the idea that the media are agents of social control who consciously and unconsciously reward conformity to and punish deviance from a perceived mainstream or status quo [@mcleodSocialControlSocial1999].
In this regard, @shoemakerMediaTreatmentDeviant1984 found that groups that are perceived as more deviant or radical tend to receive more delegitimising coverage [@hallPolicingCrisisMugging1978]^[Also see the deviance sphere in @hallinUncensoredWarMedia1986.].
Based on this, @boyleInfluenceLevelDeviance2004 tested how the *level of deviance* of protests shape their coverage.
They conclude that coverage of deviant protest conformed closer to the *protest paradigm*, conceptualised as more critical coverage, coverage employing episodic framing and coverage that is less likely to use protesters as sources.

How this might work in practice was provided by @tenenboimweinblattProducingProtestNews2014, who studied in in-depth interviews how journalists perceived their hand in framing a social protest in Israel.
At first, individual journalists were supportive of the movement as they identified strongly with protesters, who belonged to a similar socio-economic class, and their message, which criticised the same issues reporters faced in their daily lives.
Yet, journalists noted that this identification and sympathy for the protest demands collided with professional and procedural norms, which kept their reporting "objective" with less supportive coverage than some had hoped for.
As events progressed, substantial economic backlash from investors and advertisers, who felt targeted by protests, turned the table.
Suddenly economic and organisational demands became more noticeable to journalists.
After initial support, journalists subsequently ceased to portray the protests in a favourable light [see also @shultzinerjournalistsprotest2018].
This is in line with @curranPowerResponsibilityPress2007, who have argued that the historical control of the press by political actors has been replaced over time by market forces, which they believe to be even more effective in eliminating radical ideas from public discourse.

On a broader scale, @bennettindexing1990 has argued that influences like these lead to a pattern of "indexing" of the range of voices and interpretations of events "according to the range of views expressed in mainstream government debate about a given topic" (p. 106).
This means that in times of elite consensus, media will limit its discussion to the opinions expressed by the powerful.
Only in times of elite dissensus, the voices of the broader public will be heard.
Otherwise, alternative information or narratives are assumed to encounter fierce resistance or outright blockage [@bennettNoneDareCall2006].
Studies of war-time reporting, in particular, have lent strong empirical evidence to this supposition [@wolfsfeldMediaPoliticalConflict1997; @hallinDquncensoredWarDq1986].
Generally, it is thought that indexing hurts protesters' ability to explain the causes of their efforts [@andrewsMakingNewsMovement2010].

Specifically, reporting in the UK has been described as strongly centred, just as in other countries where free market ideals reign the press [@hallinComparingMediaSystems2004].
Opinions left and right of the Labour and Conservative spectrum are hardly ever supported by the press.
Therefore, protest goals outside the spectrum of "normal" political debate might hardly have a chance to gain traction through the press.
@curranPowerResponsibilityPress2007 argue that this is a result of market forces that limit the scope of opinions in the press in Britain.
In the 1970s and 1980s, for example, many radical and left-wing outlets were discontinued due to a lack of advertisement and subsequently funding.
Depletion of the market has led to a highly concentrated ownership structure on the national newspaper market:
According to a report by the @mediareformreport2021, only three companies (*News UK*, *Daily Mail Group* and *Reach*) control over 90% of national newspaper sales.

Theory-wise, the idea that deviant goals are less likely to be cast in a positive light is hence well supported.
However, the conceptualisation of deviance is less clear cut.
Interestingly, @boyleInfluenceLevelDeviance2004 conceptualise the *level of deviance* by using a group's goals as well as the form of protest.
The idea is that while street theatre, costumes, civil disobedience or clashes with police are one indicator of deviance, the degree of change sought by a protest is another [also @mcleodSocialControlSocial1999].
This means that a group is deemed deviant either because they do deviant things or because their goal is considered deviant from the mainstream opinion.
In later work, however, @boyleMeasuringLevelDeviance2009 proposed that goals and tactics should be treated as separate factors that exert their own influence on news treatment of protesters.

How extreme or deviant a goal is depends on the assessment of the degree of change sought by the group.
The level of deviance of goals is, therefore, often operationalised as three categories:
protests that seek to maintain the status quo (least deviant); seek minor reform; or want radical change to the status quo (most deviant) [@boyleInfluenceLevelDeviance2004; @boyleNewspapersProtestExamination2005; @boyleMeasuringLevelDeviance2009; @boyleAdherenceProtestParadigm2012].
The distinction of protesters' goals between support or threat to the status quo, however, is difficult, if not arbitrary, as @kilgoHierarchy2019 note:
there is no irrefutable conceptualisation of the "status quo" as political and social realities are constantly changing; a goal that was a threat to the status quo five or ten years ago could be in favour of the current situation.
@boyleMeasuringLevelDeviance2009, for example, tested this idea by comparing the treatment of pro-life and pro-choice protests before and after the ruling in the Roe v. Wade case --- which effectively legalised abortion in the entire US.
They found only partial support for the importance of the status quo for protest coverage as pro-life protesters were treated more critically after their goal became to change the status quo, but pro-choice protesters were treated the same before and after the law changed.
Similarly, @boyleNewspapersProtestExamination2005 found that while the level of deviance of most protest movements decreased over time, headline and article valence stayed roughly the same.
Analysing the deviance of protest goals from the status quo, therefore, appears less meaningful.

Besides the level of deviance, studies also assessed the type of goal as a factor for explaining coverage of protests.
The specific protest categories differ between studies:
@boyleInfluenceLevelDeviance2004 and @boyleNewspapersProtestExamination2005 provided that in a range of events between 1960 and 1999, anti-war protesters were treated most critically, followed by labour protests (e.g., strikes and pickets), protests against wrongdoings by the police [only @boyleInfluenceLevelDeviance2004] and social-issue protests (e.g., abortion or hunger) which were treated most favourably.
In an analysis of newer coverage (2007-2009), @boyleAdherenceProtestParadigm2012 found that *political protests*, by which they mean protests seeking political change, were treated more critically than anti-war and social-issue protests.

Instead of ranking these types of protest according to the level of deviance, @kilgoHierarchy2019 suggest that "journalistic routines are subject to a *hierarchy of social struggle*, in which certain topics are given precedence and legitimacy, and others are delegitimised, trivialised, or ignored altogether" (p. 16).
They also add several new types to this list --- only some of which apply to the UK.
Specifically, they found that coverage of protest against anti-Black and anti-Indigenous racism, international protest, and anti-Trump protests use frames that delegitimise their issue far more often than other protest types.
For some issues, like health, the environment and immigration protests, the press used legitimising frames more often.
Similarly, @harlowWholeWorldWatching2020 coded 33 specific types of protests before summarising them into six categories.
They find that while not all protest types had a significant influence on the use of frames, some patterns emerged:
conservative protests, for example, were less likely to be covered in a riot frame but more likely to be covered in a spectacle frame, while socio-economic protests were more likely to be covered using the riot frame.
For a legitimising frame, dubbed debate frame, no significant relationship to a protest type was found.

Following @kilgoHierarchy2019, I expect that some goals are more likely to receive legitimising and some are more likely to receive delegitimising coverage.
In other words, I expect that a *hierarchy of social struggle* will emerge in UK reporting of protest, which can predict which causes are given precedence over others.
I will use five broad categories for protest goals, ordered from most to least likely to receive critical coverage these are:
*political*, *anti-war*, *labour*, *police* and *social-issue* protests [@boyleAdherenceProtestParadigm2012; @boyleNewspapersProtestExamination2005; @boyleInfluenceLevelDeviance2004].
The hypotheses I will test are thus:

- `r hyp("H3a")`
- `r hyp("H3b")`

Note that the expected hierarchy in **H3b** is the flipped version of **H3a** as I expect that the likelihood for a protest to be covered with a legitimising frame should be roughly diametrically opposed to the likelihood for a protest to be covered with a delegitimising frame.

#### Tactics

As mentioned above, early studies often treated goals and tactics as two sides of the same coin.
However, @boyleMeasuringLevelDeviance2009 established that this is a shortcoming as both have distinct effects on the coverage of protest.
Specifically, they showed that when both are taken into account, extreme protest tactics facilitate critical coverage stronger than protest goals.
This insight had already been observed by one of the earliest studies of protest coverage:
@gitlinWholeWorldWatching1980 found that the amount of outright negative coverage, instead of trivialisation, increased as anti-war protests became more extreme.
@boyleAdherenceProtestParadigm2012, who additionally test the effects of protest location, protest type, and a group's goals in an international comparison of protest coverage, conclude that the tactics employed during a protest are the strongest determinant of delegitimising coverage.
Likewise, @WasowSeeding2020 concludes that nonviolent black-led protests between 1960 and 1972 led to more coverage of civil rights issues, while protester violence invoked language associated with disorder and calls for law and order.

How extreme the tactics are is usually determined by two variables:
do the protesters break laws and do they engage in violence?
From a broader perspective, a focus on extreme protest tactics is, therefore, in line with the professional norms of "newsworthiness" mentioned above:
usually, the clashes rather than the message of a protest will produce most conflict, drama, unusualness and negativity.
This leads to the following expectations:

- `r hyp("H3c")`
- `r hyp("H3d")`

The relation between violent tactics and coverage has been known for a long time by both researchers and protesters [@gitlinWholeWorldWatching1980].
When tactics become more violent, coverage becomes less supportive of the cause and more supportive of repressive measures.
Nevertheless, protesters often use violent tactics because this will usually lead to (more) coverage [@boyleAdherenceProtestParadigm2012].
**H3c** and **H3d**, therefore, also answer the question if this bargain makes sense for a movement.

#### State Response

As mentioned above, @WasowSeeding2020 found in his study of black-led protests between 1960 and 1972 that protester-initiated violence led to delegitimising coverage and subsequently turned public opinion and voter preferences against the civil rights movement.
However, @WasowSeeding2020 also tested the effects of violent state response towards the movement.
Specifically, he found that there are four scenarios that yielded slightly different outcomes:
When the state tolerated peaceful protest, this lead to coverage which legitimised the concerns voiced by protesters;
when the state used violent repression against peaceful protesters, this lead to coverage that is more supportive of the protesters' goal --- and to an overall higher volume of coverage about the respective event;
when only the protesters were violent, the civil rights movement was penalised in the coverage;
the same was, however, true when both police and protesters engaged in substantial violence.
In this last case, the coverage is predicted to focus on concerns about order and public safety.

Wasow's [-@WasowSeeding2020] study thus mostly confirms the relationship between violence and (de)legitimising coverage formalised in **H3c** and **H3d**.
A scenario in which the state uses violent repression against peaceful protesters, however, has not been covered by any previous studies.
Following @WasowSeeding2020, I include this additional layer, which leads to the expectation that:

- `r hyp("H3e")`

### Outlet-Level Factors

#### Newspaper Ideology

In the landmark study which defined the idea of a *protest paradigm*, @chanJournalisticParadigmsCivil1984 found that not all newspapers engage in delegitimising protests in the same way:
the left-leaning media were less hostile towards the protesting teachers and students in Hong Kong than centrist media, which were in turn less hostile than right-leaning journalists.
This factor has usually been overlooked in subsequent US-based studies, which might assume that ideological differences between major outlets are less important given the emphasis on objectivity in the country's journalistic culture.
More recent studies, however, have started again to employ newspaper ideology as an explaining factor for the level of marginalisation different protests face [@leeTriggeringProtestParadigm2014; @shahinProtestingParadigmComparative2016; @ozMainstreamMediasCoverage2016].
As mentioned above, UK media are distinctly more partisan than outlets in the US and elsewhere [@kuhnPoliticsMediaBritain2007; @hallinComparingMediaSystems2004], which is likely to increase the effect of newspaper ideology on protest reporting.
This leads to the expectation:

- `r hyp("H3f")`
- `r hyp("H3g")`

@weaverRevisitingProtestParadigm2012 add nuance to this:
more than by their right or left ideology, the coverage of an outlet might be determined by the distance of their own ideological tilt to the ideology of a protest in question.
They test this using prime-time cable news coverage about the right-leaning *Tea Party* movement in the United States in 2009-2010 to assess which outlets used marginalisation devices more often.
Their finding suggests that liberally aligned MSNBC was most likely to marginalise the *Tea Party*, followed by the centrist AP and CNN, and Fox News being supportive overall.
Since protests have traditionally held goals that were considered left more often in Western countries --- and continue to do so [@borbathDifferentWorldsContention2020] --- the assumption that right-leaning media are generally more critical of protest might be incorrect.
This leads to two hypotheses that will be tested:

- `r hyp("H3h")`
- `r hyp("H3i")`

#### Type of Newspaper

Gavin [-@gavinPressTelevisionBritish2007, pp. 95-118] and @rosieAnarchistsWorldCup2009 add another factor related to newspaper outlets that is especially pronounced in the UK:
the difference between tabloid and broadsheet newspapers.
@gavinPressTelevisionBritish2007, who assessed the May Day demonstration in London in 2000, concluded that tabloid outlets used elements from the *protest paradigm* more extensively and failed to mention any motivation or goals of the protest.
Broadsheet newspapers --- especially *The Guardian* --- apparently made an effort to explain why the protests were held.
In their qualitative study, @rosieAnarchistsWorldCup2009 find that tabloid newspapers were more extreme in their framing of peaceful G8 protesters as a fringe group of radical militants.
The sharp divide in reporting style between broadsheets newspapers (*The Daily Telegraph*, the *Financial Times*, *The Guardian*, *The Independent* and *The Times*), and tabloids (*Daily Express*, *Daily Mail*, the *Daily Mirror* and *The Sun*) is well-known and essentially unique to the UK media system.
In general, it is expected that tabloid newspapers are more focused on human interest and entertainment [@kuhnPoliticsMediaBritain2007; @hallinComparingMediaSystems2004].
It is, therefore, only plausible to expect that the two types use different frames.
Following the work by @rosieAnarchistsWorldCup2009, the following expectations can be made:

- `r hyp("H3j")`
- `r hyp("H3k")`

### Time-Bound Factors

Two factors related to time are expected to have an influence on reporting about protests.
The first one was already mentioned above:
the year in which an article was published is expected to determine if it is being framed in a delegitimising way, with newer articles using delegitimising framing less and legitimising framing more often (**H2a** and **H2b**).

The second factor very likely only affects large scale, highly mediatised protests.
@gottliebProtestNewsFraming2015 suggests that coverage follows a *news framing cycle*:
coverage of protest begins with a focus on conflict --- hence marginalising or ignoring the goals and grievances of a protest --- then shifts to more substantial coverage before returning to conflict.
This concept, based on @downsAttentionCycle1972, follows the logic that journalists will use different frames to cover an event as time passes "to keep the story alive and fresh" [@chyiMediaSalienceProcess2004, p. 22].

The *news framing cycle* intuitively makes sense, yet the proposed time frame might suffer from one problem:
Gottlieb’s [-@gottliebProtestNewsFraming2015] case, the Occupy Wall Street movement, was arguably special since protest stretched over a long period and media attention did not wane for months.
He, therefore, suggests that the second phase, when reporting shifts to more substantial coverage, started only in the third week after the protest started.
This finding is probably not suitable for generalisation as attention to most protests wanes after days, rather than weeks or months.
In most cases, the media will have forgotten about a protest the day after the first reports were printed.

Nevertheless, the idea of a protest news framing cycle seems worth testing, albeit in an abbreviated form:

- `r hyp("H3l")`
- `r hyp("H3m")`