### 2022-03-28 sage/sage.Rmd

---
title: "Whose voices are heard? Making sense of framing patterns in protest coverage"
runninghead: "Johannes B. Gruber"
author:
- name: "Johannes B. Gruber"
  num: 1
address:
- num: 1
  org: "European New School of Digital Studies (ENS), European University Viadrina Foundation Frankfurt (Oder)"
corrauth: "Johannes B. Gruber, European New School of Digital Studies (ENS), European University Viadrina Foundation Frankfurt (Oder)."
email: "Gruber@Europa-Uni.de"
abstract: "`r readLines('../paper/abstract.md')`"
keywords: protest paradigm; media framing
classoption:
  - Royal
  - times
bibliography: ../paper/references.bib
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
---

```{r article, child = "../paper/article.Rmd"}
```