### 2023-06-14 paper/app-eventdata.qmd
finished methods

## Selected and Recoded protest events {#app-eventdata .appendix}

Events were filtered from @ClarkMMPdata2019 to include only protests in the UK.
The table contains the ideology variable that was added through annotation of events.

```{r}
#| eval: false
rio::import("../misc/mmp_events_coded_table.csv") %>%
  knitr::kable(format = "latex",
               booktabs = TRUE,
               longtable = TRUE,
               linesep = "",
               caption = "Employed Subset of the MMP Data by Clark and Regan (2019) with Additional and Recoded Variables",
               label = NA) %>%
  kableExtra::kable_styling(
    full_width = FALSE,
    latex_options = "repeat_header",
    font_size = 10
  )%>% 
  kableExtra::column_spec(c(1:3, 5), width = "1.8cm") %>%
  kableExtra::column_spec(4, width = "6cm")
```